R Notebook

### 2019-07-30 src/testing_file.Rmd
feckit

```{r}
df <- iris
df <- na.omit(df)
df <- scale(df[ ,-5])

# Dissimilarity matrix
d <- dist(df, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
```

```{r}
iris2 <- pres[ ,-12]
species_labels <- pres[ ,12]
library(colorspace) # get nice colors
species_col <- rev(rainbow_hcl(4))[as.numeric(species_labels)]

# Plot a SPLOM:
pairs(iris2, col = species_col,
      lower.panel = NULL,
       cex.labels=2, pch=19, cex = 1.2)

# Add a legend
par(xpd = TRUE)
legend(x = 0.05, y = 0.4, cex = 2,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col))
par(xpd = NA)
```

```{r}
par(las = 1, mar = c(4.5, 3, 3, 2) + 0.1, cex = .8)
MASS::parcoord(iris2, col = species_col, var.label = TRUE, lwd = 2)

# Add Title
title("Parallel coordinates plot of the Iris data")
# Add a legend
par(xpd = TRUE)
legend(x = 1.75, y = -.25, cex = 1,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col), horiz = TRUE)
```

```{r}
d_iris <- dist(iris2) # method="man" # is a bit better
hc_iris <- hclust(d_iris, method = "complete")
iris_species <- rev(levels(pres[,12]))

library(dendextend)
dend <- as.dendrogram(hc_iris)
# order it the closest we can to the order of the observations:
dend <- rotate(dend, 1:6361)

# Color the branches based on the clusters:
dend <- color_branches(dend, k=4) #, groupLabels=iris_species)

# Manually match the labels, as much as possible, to the real classification of the flowers:
labels_colors(dend) <-
   rainbow_hcl(3)[sort_levels_values(
      as.numeric(pres[,12])[order.dendrogram(dend)]
   )]

# We shall add the flower type to the labels:
labels(dend) <- paste(as.character(iris[,5])[order.dendrogram(dend)],
                           "(",labels(dend),")", 
                           sep = "")
# We hang the dendrogram a bit:
dend <- hang.dendrogram(dend,hang_height=0.1)
# reduce the size of the labels:
# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
dend <- set(dend, "labels_cex", 0.5)
# And plot:
par(mar = c(3,3,3,7))
plot(dend, 
     main = "Clustered Iris data set
     (the labels give the true flower species)", 
     horiz =  TRUE,  nodePar = list(cex = .007))
legend("topleft", legend = iris_species, fill = rainbow_hcl(3))
```

```{r}
library(circlize)
par(mar = rep(0,4))
circlize_dendrogram(dend)
```

```{r}
iris3 <- pres[ ,-2]
species_labels <- pres[ ,2]
library(colorspace) # get nice colors
species_col <- rev(rainbow_hcl(12))[as.numeric(species_labels)]

# Plot a SPLOM:
pairs(iris3, col = species_col,
      lower.panel = NULL,
       cex.labels=2, pch=19, cex = 1.2)

# Add a legend
par(xpd = TRUE)
legend(x = 0.05, y = 0.4, cex = 2,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col))
par(xpd = NA)
```

```{r}
d_iris2 <- dist(pres) # method="man" # is a bit better
hc_iris2 <- hclust(d_iris2, method = "complete")
iris_species2 <- rev(levels(pres[,12]))

library(dendextend)
dend2 <- as.dendrogram(hc_iris2)
# order it the closest we can to the order of the observations:
dend2 <- rotate(dend2, 1:6361)

# Color the branches based on the clusters:
dend2 <- color_branches(dend2, k=4) #, groupLabels=iris_species)

# Manually match the labels, as much as possible, to the real classification of the flowers:
labels_colors(dend2) <-
   rainbow_hcl(3)[sort_levels_values(
      as.numeric(pres[,12])[order.dendrogram(dend2)]
   )]

# We shall add the flower type to the labels:
labels(dend2) <- paste(as.character(iris[,5])[order.dendrogram(dend2)],
                           "(",labels(dend2),")", 
                           sep = "")
# We hang the dendrogram a bit:
dend2 <- hang.dendrogram(dend2,hang_height=0.1)
# reduce the size of the labels:
# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
dend22 <- set(dend2, "labels_cex", 0.5)
# And plot:
par(mar = c(3,3,3,7))
plot(dend2, 
     main = "Clustered Iris data set
     (the labels give the true flower species)", 
     horiz =  TRUE,  nodePar = list(cex = .007))
legend("topleft", legend = species_labels, fill = rainbow_hcl(3))
```

```{r}
library(circlize)
par(mar = rep(0,4))
circlize_dendrogram(dend2)
```

@@

data_aea <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
data_aea$depthlyr_mod <- NA
data_aea$predicted <- NA
data_aea <- completeFun(data_aea, "depth_layer")

@@

```{r}
mth <- c("1", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
mthn <- c("jan", "mar", "apr", "may", "jun", "july", "aug", "sep", "oct", "nov","dec")
aea <- raster("../output/env/aea.tif")

#for (m in 1:length(mth)){
#mont <- mth[m]
mont <- 7
data_aeaX <- subset(data_aea, month == mont)
xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
for (j in 1:nrow(data_aea_sp)) { 
for (i in 1:length(avgasc_list)) {  
 de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
  mth <- (data_aea_sp$month[j])  # a variable for the observation month
  rasdep <- str_sub(avgasc_list[i], 48, -5)
       if (mth == mont & rasdep == de){
          temp_brick <- raster(avgasc_list[i])
        crs(temp_brick) <- aea@crs 
              data_aea_sp$predicted[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])

@@

write.csv(data_aea1, "../output/maxent/all_data/testforslr_jul.csv", row.names = FALSE)
 #}

@@

@@

```{r}
mod <- readRDS(file = "../output/maxent/all_data/maxmod.rds", refhook = NULL)
px <- response(mod@models[[3]])
```

>>>>

@@

data_aea <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
data_aea$depthlyr_mod <- NA
data_aea$predicted_alldata <- NA
data_aea <- completeFun(data_aea, "depth_layer")

@@

```{r}
#mth <- c("1", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
#mthn <- c("jan", "may", "apr", "may", "jun", "july", "aug", "sep", "oct", "nov","dec")
aea <- raster("../output/env/aea.tif")

#for (m in 1:length(mth)){
#mont <- mth[m]
mont <- 11
data_aeaX <- subset(data_aea, month == mont)
xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/nov/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
for (j in 1:nrow(data_aea_sp)) { 
for (i in 1:length(avgasc_list)) {  
 de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
  mth <- (data_aea_sp$month[j])  # a variable for the observation month
  rasdep <- str_sub(avgasc_list[i], 47, -5)
       if (mth == mont & rasdep == de){
          temp_brick <- raster(avgasc_list[i])
        crs(temp_brick) <- aea@crs 
              data_aea_sp$predicted[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])

@@

write.csv(data_aea1, "../output/maxent/all_data/pred_file_nov.csv", row.names = FALSE)
 #}

@@

aea <- raster("../output/env/aea.tif")

#for (m in 1:length(mth)){
#mont <- mth[m]
mont <- 11
data_aeaX <- read.csv("../output/maxent/all_data/pred_file_nov.csv", header = TRUE)
data_aeaX$mthpred <- NA
xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
avgasc_list <- list.files(paste0("../output/maxent/predictions/nov/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
for (j in 1:nrow(data_aea_sp)) { 
for (i in 1:length(avgasc_list)) {  
 de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
  mth <- (data_aea_sp$month[j])  # a variable for the observation month
  rasdep <- str_sub(avgasc_list[i], 46, -5)
       if (mth == mont & rasdep == de){
          temp_brick <- raster(avgasc_list[i])
        crs(temp_brick) <- aea@crs 
              data_aea_sp$mthpred[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])

#}
              data_aea1 <- as.data.frame(data_aea_sp)
              write.csv(data_aea1, "../output/maxent/all_data/pred_file_nov.csv", row.names = FALSE)

@@

```{r}
mod <- readRDS(file = "../output/maxent/all_data/maxmod.rds", refhook = NULL)
px <- response(mod@models[[4]])
```

#get all probabiltiy values for a period of time and find the upper 20%

for july with all data
```{r}
clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) # first list the files containing the data
a <- stack(clogasc_list) #stack the files
b <- getValues(a) # create a matrix of the values 
c <- as.vector(b) #make into vector
d <- na.omit(c) #remove all na values
d <- sort(d, decreasing = TRUE) # order from high to low
drow <- length(d) # how many values are there
dr20 <- (drow*20)/100  # calculate which value is the upper 20%
d20 <- d[dr20] # 0.149577....
dmaxv <- max(d) #0.998087
dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.998097...  = 0.798477....

#binary threshold function d20
bind20 <- function(x){
  ifelse(x >= d20, 1,
  ifelse(x < d20, 0, NA)) 
}

#binary threshold function dm20
bindm20 <- function(x){
  ifelse(x >= dm20, 1,
  ifelse(x < dm20, 0, NA)) 
}

for (cum in 1:length(clogasc_list)){
  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cum_d20 <- calc(cumasc, fun = bind20)
  cum_dm20 <- calc(cumasc, fun = bindm20)
  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/july_alldata/july_alldata_d20_", ascname, ".asc"), overwrite = TRUE)
  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/july_alldata/july_alldata_dm20_", ascname, ".asc"), overwrite = TRUE)
  }

binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_alldata_d20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/d20_2d.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/d20_2d.png"))
dev.off()
binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_alldata_dm20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/dm20_2d.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/dm20_2d.png"))
dev.off()

#make map

aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
#brk <- seq(0.1, 1, 0.1) #for legend
redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
pronly <- subset(presback, occurrence == 1)
prjul <- subset(pronly, month == 7)
#prjan <- subset(pronly, month == 1)

  pras <- raster("../output/maxent/slr_analysis/july_alldata/d20_2d.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/d20_2d.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

  pras <- raster("../output/maxent/slr_analysis/july_alldata/dm20_2d.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/dm20_2d.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

for july with just july data
```{r}
clogasc_list <- list.files(paste0("../output/maxent/predictions/jul/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) # first list the files containing the data
a <- stack(clogasc_list) #stack the files
b <- getValues(a) # create a matrix of the values 
c <- as.vector(b) #make into vector
d <- na.omit(c) #remove all na values
d <- sort(d, decreasing = TRUE) # order from high to low
drow <- length(d) # how many values are there
dr20 <- (drow*20)/100  # calculate which value is the upper 20%
d20 <- d[dr20] # 0.083319....
dmaxv <- max(d) #0.885157....
dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.885157....  = 0.70812....

#binary threshold function d20
bind20 <- function(x){
  ifelse(x >= d20, 1,
  ifelse(x < d20, 0, NA)) 
}

#binary threshold function dm20
bindm20 <- function(x){
  ifelse(x >= dm20, 1,
  ifelse(x < dm20, 0, NA)) 
}

for (cum in 1:length(clogasc_list)){
  cumascdep <- str_sub(clogasc_list[cum], start = 46, end = -5) #selects the depth from the file name
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cum_d20 <- calc(cumasc, fun = bind20)
  cum_dm20 <- calc(cumasc, fun = bindm20)
  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/july_alldata/july_monthdata_d20_", ascname, ".asc"), overwrite = TRUE)
  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/july_alldata/july_monthdata_dm20_", ascname, ".asc"), overwrite = TRUE)
  }

binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_monthdata_d20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/d20_2d_month.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/d20_2d_month.png"))
dev.off()
binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_monthdata_dm20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/dm20_2d_month.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.png"))
dev.off()

#make map

aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
#brk <- seq(0.1, 1, 0.1) #for legend
redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
pronly <- subset(presback, occurrence == 1)
prjul <- subset(pronly, month == 7)
#prjan <- subset(pronly, month == 1)

  pras <- raster("../output/maxent/slr_analysis/july_alldata/d20_2d_month.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/d20_2d_month.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

  pras <- raster("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))
```

for jan with all data
```{r}
clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/jan/"), pattern = "^avg_.*.asc", full.names = TRUE) # first list the files containing the data
a <- stack(clogasc_list) #stack the files
b <- getValues(a) # create a matrix of the values 
c <- as.vector(b) #make into vector
d <- na.omit(c) #remove all na values
d <- sort(d, decreasing = TRUE) # order from high to low
drow <- length(d) # how many values are there
dr20 <- (drow*20)/100  # calculate which value is the upper 20%
d20 <- d[dr20] # 0.08307....
dmaxv <- max(d) #0.996738...
dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.996738...  = 0.7973906....

#binary threshold function d20
bind20 <- function(x){
  ifelse(x >= d20, 1,
  ifelse(x < d20, 0, NA)) 
}

#binary threshold function dm20
bindm20 <- function(x){
  ifelse(x >= dm20, 1,
  ifelse(x < dm20, 0, NA)) 
}

for (cum in 1:length(clogasc_list)){
  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cum_d20 <- calc(cumasc, fun = bind20)
  cum_dm20 <- calc(cumasc, fun = bindm20)
  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_alldata_d20_", ascname, ".asc"), overwrite = TRUE)
  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_alldata_dm20_", ascname, ".asc"), overwrite = TRUE)
  }

binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_alldata_d20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/d20_2d.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/d20_2d.png"))
dev.off()
binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_alldata_dm20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/dm20_2d.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/dm20_2d.png"))
dev.off()

#make map

aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
#brk <- seq(0.1, 1, 0.1) #for legend
redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
pronly <- subset(presback, occurrence == 1)
#prjul <- subset(pronly, month == 7)
prjan <- subset(pronly, month == 1)

  pras <- raster("../output/maxent/slr_analysis/jan_alldata/d20_2d.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/d20_2d.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

  pras <- raster("../output/maxent/slr_analysis/jan_alldata/dm20_2d.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/dm20_2d.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

for jan with just jan data
```{r}
clogasc_list <- list.files(paste0("../output/maxent/predictions/jan/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) # first list the files containing the data
a <- stack(clogasc_list) #stack the files
b <- getValues(a) # create a matrix of the values 
c <- as.vector(b) #make into vector
d <- na.omit(c) #remove all na values
d <- sort(d, decreasing = TRUE) # order from high to low
drow <- length(d) # how many values are there
dr20 <- (drow*20)/100  # calculate which value is the upper 20%
d20 <- d[dr20] # 0.01165....
dmaxv <- max(d) #0.73946....
dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.73946....  = 0.591569....

#binary threshold function d20
bind20 <- function(x){
  ifelse(x >= d20, 1,
  ifelse(x < d20, 0, NA)) 
}

#binary threshold function dm20
bindm20 <- function(x){
  ifelse(x >= dm20, 1,
  ifelse(x < dm20, 0, NA)) 
}

for (cum in 1:length(clogasc_list)){
  cumascdep <- str_sub(clogasc_list[cum], start = 46, end = -5) #selects the depth from the file name
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cum_d20 <- calc(cumasc, fun = bind20)
  cum_dm20 <- calc(cumasc, fun = bindm20)
  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_monthdata_d20_", ascname, ".asc"), overwrite = TRUE)
  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_monthdata_dm20_", ascname, ".asc"), overwrite = TRUE)
  }

binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_monthdata_d20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/d20_2d_month.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.png"))
dev.off()
binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_monthdata_dm20_.*.asc", full.names = TRUE)
binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.asc", overwrite=TRUE)
plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.png"))
dev.off()

#make map

aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
#brk <- seq(0.1, 1, 0.1) #for legend
redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
pronly <- subset(presback, occurrence == 1)
prjul <- subset(pronly, month == 7)
#prjan <- subset(pronly, month == 1)

  pras <- raster("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))

  pras <- raster("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.asc")
  crs(pras) <- apr
  ascname <- names(pras)
  pras[pras >0] <- 1
  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
  print(lp)
  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.png"), width = 1920,height = 1080)
  dev.off() # stops automatic saving of the plot to a png))
```

mont <- 
data_aeaX <- read.csv("../output/maxent/all_data/testforslr_jan.csv", header = TRUE)
data_aeaX$predicted_mthdata <- NA

xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
avgasc_list <- list.files(paste0("../output/maxent/predictions/jan"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
for (j in 1:nrow(data_aea_sp)) { 
for (i in 1:length(avgasc_list)) {  
 de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
  rasdep <- str_sub(avgasc_list[i], 46, -5)
       if (rasdep == de){
          temp_brick <- raster(avgasc_list[i])
        crs(temp_brick) <- aea@crs 
              data_aea_sp$predicted_mthdata[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])

           data_aea1 <- as.data.frame(data_aea_sp)

write.csv(data_aea1, "../output/maxent/all_data/predictions_jan.csv", row.names = FALSE)
 #}

#code to create benthic layers

```{r}
aea <- raster("../output/env/aea.tif") #the correct CRS
benlst <- list.files("../data/env/ascii/benthic_layers/template", pattern = ".*.asc", full.names = TRUE)
apr <- aea@crs

for (pred in 1:length(benlst)){
  pras <- raster(benlst[pred])
  crs(pras) <- apr
  ascname <- names(pras)
  depth <-  (sapply(strsplit(ascname, "_"), "[[", 4)) # extracting the third part of the netcdf (inc.nc)
  pras[pras >0] <- 0
  pras[is.na(pras)] <- 1
  writeRaster(pras, paste0("../data/env/ascii/benthic_layers/done/benthic_", depth, ".asc"), overwrite = TRUE)

#just check it works
benlst <- list.files("../data/env/ascii/benthic_layers/done", pattern = ".*.asc", full.names = TRUE)

for (pred in 1:length(benlst)){
  pras <- raster(benlst[pred])
  plot(pras)
}

```{r}
predjan <- read.csv("../output/maxent/all_data/predictions_jan.csv", header = TRUE)
plot(predjan$occurrence, predjan$predicted_mthdata)
plot(predjan$occurrence, predjan$predicted_alldata)
```

```{r}
predjul <- read.csv("../output/maxent/all_data/predictions_jul.csv", header = TRUE)
plot(predjul$occurrence, predjul$predicted_mthdata)
plot(predjul$occurrence, predjul$predicted_alldata)
```

```{r}
predjan <- read.csv("../output/maxent/all_data/predictions_jan.csv", header = TRUE)

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

predjan <- completeFun(predjan, "predicted_mthdata")
predjanoc <- subset(predjan, occurrence == 1)
predjanbk <- subset(predjan, occurrence == 0)
noc <- nrow(predjanoc)
nbk <- nrow(predjanbk)
janmth5plusoc <- sum(predjanoc$predicted_mthdata > 0.5)
janmth5plusbk <- sum(predjanoc$predicted_mthdata > 0.5)
janmthpp <- (janmth5plusoc/noc)*100
janmthbkp <- (janmth5plusbk/nbk)*100

janall5plusoc <- sum(predjanoc$predicted_alldata > 0.5)
janall5plusbk <- sum(predjanoc$predicted_alldata > 0.5)
janallpp <- (janall5plusoc/noc)*100
janallbkp <- (janall5plusbk/nbk)*100
```

```{r}
hist(predjanoc$predicted_mthdata)
hist(predjanoc$predicted_alldata)
hist(predjanbk$predicted_mthdata)
hist(predjanbk$predicted_alldata)
```
